Structural controllability of ecological networks

### 2018-07-17 paper/manuscript.Rmd
Feedback Daniel

1. An important dimension of a species role is its ability to alter the state and maintain the diversity of its community. 
Centrality metrics have often been used to identify these species, which are sometimes referred as "keystone" species. 
However, the relationship between centrality and keystoneness is phenomenological and largely based on an intuitive understanding of what may constitute an important species. 
While centrality is good at predicting which species could cause the largest change in their community it says little about how these species could be used to attain (or preserve) a particular state of the community.

2. Here we introduce structural controllability an approach that allows us to quantify the extent to which network topology can be harnessed to achieve a desired state. 
It also allows us to quantify a species control capacity—its relative importance—and identify the set of species that, collectively, are critical in this context.
We illustrate the application of structural controllability with ten pairs of uninvaded and invaded plant-pollinator communities.

3. We found that the controllability of a community is not dependent on its invasion status, but on the asymmetric nature of its mutual dependences. 
While central species were also likely to have a large control capacity, centrality fails to identify species that despite being less connected were critical in their communities. 
Interestingly, this set of critical species was mostly composed of plants and every invasive species in our dataset was part of it. 
We also found that species with high control capacity and in particular critical species contribute the most to the to stable coexistence of their community even when controlling for its degree, abundance/interaction strength, and the relative dependence of their partners.

@@

**Keywords:** Keystone species, ecological networks, management interventions, mutualism, network control theory, structural controllability, plant population and community dynamics, species importance

>>>>

1. An important dimension of a species role is its ability to alter the state and maintain the diversity of its community. 
Centrality metrics have often been used to identify these species, which are sometimes referred as "keystone" species. 
However, the relationship between centrality and keystoneness is largely phenomenological and based mostly on our intuition regarding what constitutes an important species. 
While centrality is useful when predicting which species extinctions could cause the largest change in a community, it says little about how these species could be used to attain or preserve a particular community state.

2. Here we introduce structural controllability, an approach that allows us to quantify the extent to which network topology can be harnessed to achieve a desired state. 
It also allows us to quantify a species control capacity—its relative importance—and identify the set of species that, collectively, are critical in this context.
We illustrate the application of structural controllability with ten pairs of uninvaded and invaded plant-pollinator communities.

3. We found that the controllability of a community is not dependent on its invasion status, but on the asymmetric nature of its mutual dependences. 
While central species were also likely to have a large control capacity, centrality fails to identify species that despite being less connected were critical in their communities. 
Interestingly, this set of critical species was mostly composed of plants and every invasive species in our dataset was part of it. 
We also found that species with high control capacity and in particular critical species contribute the most to the stable coexistence of their community.
This result was true, even when controlling for its degree, abundance/interaction strength, and the relative dependence of their partners.

@@

**Keywords:** Keystone species, management interventions, mutualism, network control theory, plant population and community dynamics, species importance, control capacity
